---
title: "Mapping ex report"
author: "Jiaheng Li"
date: "2020/11/11"
output: 
  revealjs::revealjs_presentation:
    self_contained: false
    reveal_plugins: ["search", "zoom", "notes", "chalkboard", "menu"]
    reveal_options:
      chalkboard: 
        theme: chalkboard
        toggleNotesButton: true
      menu:
        numbers: true
---
## Slide with Mapping ex report

- Project Description 
- Data Cleaning 
- Using template on Shiny website 

## 1 Project Description

In this project, I am using the OpenFEMA Dataset: Public Assistance Funded Projects from the FEMA website. There are total 22 variables in the original dataset. There are their descriptions.

disasterNumber: Sequentially assigned number used to designate an event or incident declared as a disaster. For more information on the disaster process visit https://www.fema.gov
pwNumber: Sequentially assigned number used to identify a unique project
applicationTitle: Standard, non-unique (free form text) application title
applicantId: Unique Public Assistance applicant identification number
county: The name of a U.S. county, parish, borough, independent city or other political subdivision of a U.S. state or territory
countyCode: Unique identifier for a U.S. county; typically, this identifier makes up the last three digits of a five digit Federal Information Processing Standard (FIPS) Code
stateNumberCode:  Unique identifier for a U.S. state or territory; this identifier makes up the first two digits of a five digit Federal Information Processing Standard (FIPS) Code
state: The name of a U.S. state or territory
projectAmount: The estimated total cost of the Public Assistance grant project in dollars, without administrative costs. This amount is based on the damage survey
stateCode: Two-character unique identifier for a U.S. state or territory
totalObligated: 	The federal share of the Public Assistance grant eligible project amount in dollars, plus grantee (State) and sub-grantee (applicant) administrative costs. The federal share is typically 75% of the total cost of the project
federalShareObligated: The Public Assistance grant funding available to the grantee (State) in dollars, for sub-grantee's approved Project Worksheets
damageCategory: The category code of the damage location
dcc: Damage category code
lastRefresh: Date the record was last updated in the API data store
id: Unique ID assigned to the record
hash: MD5 Hash of the fields and values of the record
projectSize: Projects are designated as Large or Small, which can affect a number of variables in case management (processing). Project size is determined by the eligible amount as set in the damage survey
damageCategoryCode: 	Code representing the type of work eligible for reimbursement through a Public Assistance grant; for a full listing of each category please go to /government/grant/pa/re_categories.shtm
declarationDate: 	Date the disaster was declared
incidentType: Type of incident such as fire or flood. The incident type will affect the types of assistance available. For more information on incident types, please visit: https://www.fema.gov
obligatedDate: 	Date the grant was obligated


My purpose on this porject is to use the data I got to make a mapping tool of showing the total Obligated fund for the hurricane disaster in each state on the date when the disaster was declared.


## 2 Data Cleaning

```{r}
# data <- read_excel("C:/Users/superavis/Documents/TEMPUSE/063-superzip-example/hurricane.xlsx")
# data2 <- data %>% select(state,stateCode,totalObligated,obligatedDate)
# data2$date <- as.Date(substr(data2$obligatedDate,1,10))
# 
# data3 <- data2 %>%  group_by(date, state) %>% summarise(total  = sum(totalObligated,na.rm=T))
# save(data3, file = "hurricane.rdata")
```

I firstly use the excel filter to filter out the hurricane data in the original data because the size of the original data was too large for R Studio to process. Then I convert the original date to the date variable of the R studio.


## 3 Using template on Shiny website

I found a teplate on the Shiny website which can fit in the project I planned to do of the mapping assignment. (https://shiny.rstudio.com/gallery/superzip-example.html)


## 3.1 Server

Then I switched out the varibales in the template and made a new interactive map with  two parameters of 'date' and 'state' also set the output as the sum of totalObligated on that day. The codes are as following.

(https://shiny.rstudio.com/gallery/superzip-example.html)

```{r eval = FALSE} 
library(geojsonio)
library(leaflet)
library(shiny)
library(RColorBrewer)
library(scales)
library(lattice)
library(dplyr)
library(lubridate)

states <- 
  geojson_read( 
    x = "https://raw.githubusercontent.com/PublicaMundi/MappingAPI/master/data/geojson/us-states.json"
    , what = "sp"
  )


shinyServer( 
function(input, output, session) {

  labels <- reactive( { 
    
    states2 <- states
    hurr2 <- hurr[hurr$date == input$date,]
    states2$hurrican <-  hurr2$total[match(states$name, hurr2$state)]
    sprintf(
    "<strong>%s</strong><br/>%g sum of totalObligated",
    states2$name, states2$hurrican
  ) %>% lapply(htmltools::HTML) })
  
  statesdata <- reactive( {
  data.frame( hurr[hurr$date == input$date & hurr$state == input$state,3])
  })
  
  statesd <- reactive( 
  
  {
    
    states2 <- states
    hurr2 <- hurr[hurr$date == input$date,]
    
    states2$hurrican <-  hurr2$total[match(states$name, hurr2$state)]
    states2
  })
  
  output$map <-  renderLeaflet( {
  #  states2 <- statesd()
    states2 <-  states
    states2$hurrican <-  hurr$total[match(states2$name, hurr$state)]
    bins <- c(0, 10000, 50000, 1000000,Inf)
    
    pal <- colorBin("YlOrRd", domain = states2$hurrican, bins = bins)
    
    labels <- 
      sprintf(
        "<strong>%s</strong><br/>%g sum of totalObligated",
        states2$name, states2$hurrican
      ) %>% lapply(htmltools::HTML) 
  leaflet( states2) %>%
    setView(-96, 37.8, 4) %>%
    addProviderTiles("MapBox", options = providerTileOptions(
      id = "mapbox.light",
      accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN'))) %>%
    addPolygons(
      fillColor = ~pal(hurrican),
      weight = 2,
      opacity = 1,
      color = "white",
      dashArray = "3",
      fillOpacity = 0.7,
      highlight = highlightOptions(
        weight = 5,
        color = "#666",
        dashArray = "",
        fillOpacity = 0.7,
        bringToFront = TRUE),
      label = labels,
      labelOptions = labelOptions(
        style = list("font-weight" = "normal", padding = "3px 8px"),
        textsize = "15px",
        direction = "auto")) %>%
    addLegend(pal = pal, values = ~hurrican, opacity = 0.7, title = NULL,
              position = "bottomright")
  
  })
  output$tbl <- renderTable({
    statesdata()
  })


  
}
)
```

## 3.2 UI

The UI file is used for designing the layout of the mapping tool.

```{r eval = FALSE}

load("C:/Users/aaron/OneDrive/Desktop/mapping/Mapping ex/hurricane.rdata")
hurr <- data3

# Choices for drop-downs
vars1 <- as.character(hurr$date)
vars2 <- as.character(unique(hurr$state))

shinyUI( 
navbarPage("Interactive map", id="nav",

  tabPanel("U.S states hurrican distribution shinyapp",
   
        sidebarLayout( 
          sidebarPanel( 
        h2("Hurrican distribution"),

        selectInput("date", "Date", vars1, selected  = vars1[1]),
        selectInput("state", "State", vars2, selected  = vars2[1]),
        h3("sum of total totalObligated"),
        tableOutput("tbl")),
        mainPanel( 
        
        leafletOutput("map")
        )
      )
)
)
)


```

